---
title: "Step Data"
output: html_document
---

We'll first read in the data.


```{r}
activity <- read.csv("activity.csv")
summary(activity)
```



The histogram below shows the frequency of the total number of steps taken over each day.
```{r, }
sumTotalSteps <- aggregate(activity$steps, list(activity$date),sum )
colnames(sumTotalSteps) = list("Date","Total.Steps")

hist(sumTotalSteps$Total.Steps, ylim=c(0,30))

```



### What is mean total number of steps taken per day?
Here we can see the mean followed by the median of the steps taken per day.


```{r, }
mean(sumTotalSteps$Total.Steps, na.rm=TRUE)


median(sumTotalSteps$Total.Steps, na.rm=TRUE)
```

###What are the daily patterns?
Here it is determined what intervals have the highest number of steps taken over all days.
The max mean steps is early in the morning at 835.
```{r}
meanIntervalSteps <- aggregate(activity$steps, list(activity$interval),mean , na.rm=TRUE)
colnames(meanIntervalSteps) <- list("Interval","mean.steps")
plot(meanIntervalSteps$Interval, meanIntervalSteps$mean.steps, type="l")
max.AveSteps <- meanIntervalSteps[meanIntervalSteps$mean.steps == max(meanIntervalSteps$mean.steps), ]
max.AveSteps
abline(h=max.AveSteps$mean.steps, v=max.AveSteps$Interval)
```

There are a large number of missing values

```{r}

sum(is.na(activity$steps))

```

Create a new dataset with missing vales replaced by the average for that interval.


```{r}
imputed <- activity
missing_steps <- is.na(imputed$steps)
ids <- match(imputed[missing_steps, "interval"], meanIntervalSteps$Interval)
imputed[missing_steps,"steps"] <- meanIntervalSteps[ids,"mean.steps"]

```

Create a histogram using dataset with imputed missing steps.  
Frequency of 10,000-15,000 steps a day has increased
```{r}
sumTotalSteps_imputed <- aggregate(imputed$steps, list(imputed$date),sum )
colnames(sumTotalSteps_imputed) = list("Date","Total.Steps")

hist(sumTotalSteps_imputed$Total.Steps, ylim=c(0,40))
```


The mean and median haven't changed much.  Imputing using the average for that interval has more of an effect on the frequency of 10,000-15,000 steps a day.
```{r}


mean(sumTotalSteps_imputed$Total.Steps, na.rm=TRUE)
median(sumTotalSteps_imputed$Total.Steps, na.rm=TRUE)

```

### Weekday vs Weekends
The weekday has a higher peek but weekends seem to have more intervals with higher amounts of steps

```{r}
library(ggplot2)
imputed$date_type <- ifelse(weekdays(as.Date(imputed$date)) %in% c("Saturday","Sunday"), "weekend","weekday")

meanIntervalSteps_imputed <- aggregate(imputed$steps, list(imputed$date_type,imputed$interval),mean , na.rm=TRUE)
colnames(meanIntervalSteps_imputed) <- list("date_type","Interval","mean.steps")
qplot(Interval, mean.steps,data = meanIntervalSteps_imputed, factes=c(meanIntervalSteps_imputed$date_type), geom="line") +
    facet_wrap(~date_type,2)

```